---
title: "Tutorial: Controls dependent on data"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    includes:
      before_body: [includes/include_header.html, includes/include_header_navpage.html]
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# What does this tutorial achieve?

This tutorial covers the fairly advanced topic of designing controls for your shiny app that either:

- are dependent on values from datasets
- are dependent on other input values from the same Shiny app

# Our Example App

Our example app makes use of two data packages:

- `library(gapminder)`: provides socioeconomic data on 142 countries from the www.gapminder.org project
- `library(rnaturalearthhires)`: provides high resolution country borders for mapping, note that this package is not available on CRAN and must be installed using `devtools::install_github("ropenscilabs/rnaturalearthhires")`. 

Users of the app are able to view the country and country details by first selecting a continent and then a country, so both controls are dependent on the datasets and the first control subsets the number of available options for the second.

<iframe src="https://trainingidn.shinyapps.io/oxShef_shiny_tutorials_controls-depedendent-on-data/" width="100%" height="650px"></iframe>

The code for this Shiny app is on [the Github repo](https://github.com/OxShef/oxshef_shiny/tree/master/tutorial-apps/controls-depedendent-on-data) and the app is embedded into this page using a simple iframe, [see here for more details](deploy_shinyapps-io.html).

This tutorial does not walk you through building the app, instead follow these instructions:

1. [Download the shiny app code](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/OxShef/oxshef_shiny/tree/master/tutorial-apps/controls-dependent-on-data).
1. Open the controls-dependent-on-data.Rproj file from the .zip.
1. Open the server.R and ui.R files in RStudio and refer to the notes below to understand how the app works.


# updateSelectInput

The `updateSelectInput` function allows the `server.R` function to re-write the contents of a `selectInput` using information from both the `input` and `output` objects. Let's look at both instances of the `updateSelectInput` function in our template app:

1. selected_continent

```{r, eval=FALSE}
observe({
  updateSelectInput(
    session,
    "selected_continent",
    choices = unique(gapminder_countries$continent)
  )
})
```

Here the `selectInput` with the `inputID = "selected_continent"` is updated with the unique values from `gapminder_countries$continent`. The function has been wrapped in `observe` but in this instance its kind of unnecessary, as the expression is only ever going to be evaluated once (when the app loads).

2. selected_country

```{r, eval=FALSE}
observe({
  updateSelectInput(
    session,
    "selected_country",
    choices = gapminder_countries %>%
      filter(continent == input$selected_continent) %>%
      select(name) %>%
      .[[1]]
  )
})
```

Here the `observe` **is** necessary as the `selectInput` with the `inputId = "selected_country"` needs to be updated whenever the `input$selected_continent` variable is changed. The `observe` function is *tickled* by any `input$` variable inside of it, and so the list of available countries will be updated when the selected continent changes.

# Values before loading finishes

When a Shiny app first loads everything from the `ui.R` file is displayed on screen and then the `server.R` function is called with all *default input values*. This means for a very short amount of time the `selectInputs` have the values assigned in the `ui.R` file, which are as follows:

```{r, eval=FALSE}
fluidPage(
  ...,
  selectInput("serverside_continent",
                  "Selected continent:",
                  choices = ""),
  ...
selectInput("selected_country",
                  "Selected country:",
                  choices = ""),
  ...
)
```

To prevent `renderLeaflet` from briefly displaying an error because of `input$selected_country == ""` we add the following line to our `server.R` file:

```{r, eval=FALSE}
if (input$selected_country == "") {
  return()
}
```

# Can we improve this app?

The steps followed above allow for the country list to be modified by the continent selector, and for the country list to be dependent on the `gapminder` dataset. So in theory, if additional countries were added to the package in the future our app would update nicely.

However, there's still something we could do to improve our app. At the moment, each time `input$selected_country` is changed the entire map is regenerated - the "base map" is reloaded. This is often inefficient if your map contains multiple datasets and only one is changing. Fortunately, it is possible to update the `output$selected_country_map` object via the `leafletProxy` function.

The code below could be substituted for the current implementation of `output$selected_country_map`. For this toy app it doesn't massively improve the experience for the end user, but this gives you a taste of how to design more complicated apps. 

```{r, eval=F}
  output$selected_country_map <- renderLeaflet({
    if (input$selected_country == "") {
      return()
    }

    leaflet() %>%
      addTiles()

  })

  observeEvent(input$selected_country,
               {
                 if (input$selected_country == "") {
                   return()
                 }

                 country_sf <- gapminder_countries %>%
                   filter(name == input$selected_country)

                 country_bbox <- gapminder_countries %>%
                   filter(name == input$selected_country) %>%
                   st_bbox() %>%
                   as.list()

                 leafletProxy("selected_country_map") %>%
                   clearShapes() %>%
                   addPolygons(
                     data = country_sf,
                     label = ~ name,
                     popup = ~ paste("Country:", name,
                                     "<br>",
                                     "Population:", pop)
                   ) %>%
                   fitBounds(country_bbox$xmin,
                             country_bbox$ymin,
                             country_bbox$xmax,
                             country_bbox$ymax)

               })
```











