---
title: "Tutorial: Controls dependent on data"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    includes:
      before_body: [includes/include_header.html, includes/include_header_navpage.html]
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# What does this tutorial achieve?

This tutorial covers the fairly advanced topic of designing controls for your shiny app that either:

- are dependent on values from datasets
- are dependent on other input values from the same Shiny app

# Our Example App

Our example app makes use of two data packages:

- `library(gapminder)`: provides socioeconomic data on 142 countries from the www.gapminder.org project
- `devtools::install_github("ropenscilabs/rnaturalearthhires")`: provides high resolution country borders for mapping

Users of the app are able to view the country and country details by first selecting a continent and then a country, so both controls are dependent on the datasets and the first control subsets the number of available options for the second.

<iframe src="https://trainingidn.shinyapps.io/oxShef_shiny_tutorials_controls-depedendent-on-data/" width="100%" height="400px"></iframe>

The code for this Shiny app is on [the Github repo](https://github.com/OxShef/oxshef_shiny/tree/master/tutorial-apps/controls-depedendent-on-data) and the Shiny app is embedded into this page using a simple iframe, [see here for more details](deploy_shinyapps-io.html). 

# updateSelectInput

The `updateSelectInput` function allows the `server.R` function to re-write the contents of a `selectInput` using information from both the `input` and `output` objects. Let's look at both instances of the `updateSelectInput` function in our template app:

1. selected_continent

```{r, eval=FALSE}
observe({
  updateSelectInput(
    session,
    "selected_continent",
    choices = unique(gapminder_countries$continent)
  )
})
```

Here the `selectInput` with the `inputID = "selected_continent"` is updated with the unique values from `gapminder_countries$continent`. The function has been wrapped in `observe` but in this instance its kind of unnecessary, as the expression is only ever going to be evaluated once (when the app loads).

2. selected_country

```{r, eval=FALSE}
observe({
  updateSelectInput(
    session,
    "selected_country",
    choices = gapminder_countries %>%
      filter(continent == input$selected_continent) %>%
      select(name) %>%
      .[[1]]
  )
})
```

Here the `observe` **is** necessary as the `selectInput` with the `inputId = "selected_country"` needs to be updated whenever the `input$selected_continent` variable is changed. The `observe` function is *tickled* by any `input$` variable inside of it, and so the list of available countries will be updated when the selected continent changes.

# Values before loading finishes

When a Shiny app first loads everything from the `ui.R` file is displayed on screen and then the `server.R` function is called with all *default input values*. This means for a very short amount of time the `selectInputs` have the values assigned in the `ui.R` file, which are as follows:

```{r, eval=FALSE}
fluidPage(
  ...,
  selectInput("serverside_continent",
                  "Selected continent:",
                  choices = ""),
  ...
selectInput("selected_country",
                  "Selected country:",
                  choices = ""),
  ...
)
```

To prevent `renderLeaflet` from briefly displaying an error because of `input$selected_country == ""` we add the following line to our `server.R` file:

```{r, eval=FALSE}
if (input$selected_country == "") {
  return()
}
```

